---
title: "Intesity of m6A-Tracer around Ki67-marked nucleoli"
author: "Tom van Schaik"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---


### Introduction

The goal of this analysis is to determine the enrichment of m6A-Tracer signal
around nucleoli (marked by Ki67), following Ki67 pA-DamID. Dam-only signal is 
the control.

This is for RPE cells, with and without synchronization. There is only a 
Dam-only control for unsynchronized cells.


### Setup

Load the required libraries.

```{r setup, message = F, warning = F}

# Load libraries
library(ggplot2)
library(tidyverse)
library(ggbeeswarm)
library(pixmap)

# Input data
results_dir <- c("../ts201102_E1349_confocal_synchronization/ts201102_RPE_m6ATracer/ts201102_RPE_timecourse_analysis/",
                 #"../ts201207_Exxx_confocal_various_experiments/ts210316_RPE_m6ATracer/ts201207_RPE_synchronization_pADamID_analysis/",
                 "../ts210308_E1499_confocal/ts210316_RPE_m6ATracer/ts210308_RPE_m6ATracer_analysis/")

# Prepare output directory
output_dir <- "ts210615_nucleolar_enrichment"
dir.create(output_dir, showWarnings = FALSE)

# Prepare output files

```

```{r knits setup}
library(knitr)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(output_dir, "figures/"),
               cache = T)
pdf.options(useDingbats = FALSE)
```

```{r functions}

ReadPGN <- function(file.name) {
  
  # Get image, only grey values
  suppressWarnings(read.pnm(file.name)@grey*255)
  
}

```


### Strategy

...


### 1) Prepare metadata

List the different samples.

```{r prepare metadata}

# List and organize all files
metadata <- tibble(cell_statistics = dir(results_dir, recursive = T, 
                                         full.names = T,
                                         pattern = "cells_statistics")) %>%
  mutate(sample = str_remove(basename(cell_statistics),
                             "_cells_.*"),
         sample = str_remove(sample, "r._"),
         sample = str_replace(sample, "RPE", "unsync")) %>%
  separate(sample, remove = F,
           c("condition", "target", "slide")) %>%
  mutate(replicate = case_when(grepl("r1_", cell_statistics) ~ "r2",
                               grepl("r2_", cell_statistics) ~ "r3",
                               T ~ "r1"),
         replicate = factor(replicate, levels = c("r1", "r2", "r3")),
         cell = factor("RPE", levels = c("RPE")),
         condition = factor(condition, levels = c("unsync", "t0h", "t1h", "t3h",
                                                  "t6h", "t10h", "t21h")),
         target = factor(target, levels = c("Dam", "Ki67")),
         sample = paste(condition, target, replicate, slide, sep = "_")) %>%
  mutate(objects = str_replace(cell_statistics, 
                               "cells_statistics.csv",
                               "dapi_segment.pgm"),
         nucleolus_internal = str_replace(objects, 
                                          "dapi_segment", 
                                          "nucleolus_internal"),
         nucleolus_external = str_replace(objects, 
                                          "dapi_segment", 
                                          "nucleolus_external"),
         nucleolus = str_replace(objects, 
                                 "dapi_segment", 
                                 "nucleolus_smooth"),
         tracer = str_replace(objects, 
                              "dapi_segment", 
                              "m6ATracer_smooth")) %>%
  arrange(cell, target, condition, replicate, slide) %>%
  filter(condition != "t21h")

# Print metadata
metadata %>% 
  print(n = Inf)

```


### 2) Load cell sizes and create a cell table

Load the cells and show the sizes. Possibly filter on those.

```{r load cells, message = F, warning = F}

# Load all cells
LoadCells <- function(i) {
  tib <- read_csv(metadata$cell_statistics[i]) %>%
    dplyr::select(c(1:2)) %>%
    rename_all(~c("volume", "surface")) %>%
    add_column(object = 1:nrow(.),
               sample = metadata$sample[i],
               replicate = metadata$replicate[i],
               condition = metadata$condition[i],
               cell = metadata$cell[i],
               target = metadata$target[i],
               slide = metadata$slide[i]) %>%
    mutate(cell_id = paste(sample, object, sep = "_"))
  
  tib
  
}

tib_cells <- lapply(1:nrow(metadata), LoadCells) %>%
  purrr::reduce(bind_rows)

```

Plot and filter these cells.

```{r plot cells, fig.width = 5, fig.height = 3.5}

# Set size limits
limits_size <- c(10, 125)

# Plot cell volume
tib_cells %>%
  ggplot(aes(x = target, y = volume, col = replicate)) +
  geom_hline(yintercept = 0, alpha = 0) +
  geom_hline(yintercept = limits_size, col = "black", linetype = "dashed") +
  geom_quasirandom() +
  geom_boxplot(fill = NA, col = "black", outlier.shape = NA) +
  facet_grid(. ~ condition, scales = "free", space = "free") +
  xlab("") +
  ylab("Volume slice (micron)") +
  theme_bw() +
  theme(#aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Filter cells for cutoff
tib_cells <- tib_cells %>%
  mutate(size_filter = volume >= limits_size[1] & 
           volume <= limits_size[2])

tib_cells %>%
  filter(size_filter == T) %>%
  ggplot(aes(x = target, y = volume, col = replicate)) +
  geom_hline(yintercept = 0, alpha = 0) +
  geom_hline(yintercept = limits_size, col = "black", linetype = "dashed") +
  geom_quasirandom() +
  geom_boxplot(fill = NA, col = "black", outlier.shape = NA) +
  facet_grid(. ~ condition, scales = "free", space = "free") +
  xlab("") +
  ylab("Volume slice (micron)") +
  theme_bw() +
  theme(#aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


### 3) Load the pgm maps

Load the .pgm files for the various channels.

```{r load pgm, message = F, warning = F}

# Load all cells
LoadPGMMaps <- function(i, n_min = 5) {
  
  # Load objects
  objects <- ReadPGN(metadata$objects[i])
  nucleolus <- ReadPGN(metadata$nucleolus[i])
  nucleolus_internal <- ReadPGN(metadata$nucleolus_internal[i])
  nucleolus_external <- ReadPGN(metadata$nucleolus_external[i])
  tracer <- ReadPGN(metadata$tracer[i])
  
  tib <- tibble(object = c(objects),
                nucleolus_internal = c(nucleolus_internal),
                nucleolus_external = c(nucleolus_external),
                nucleolus = c(nucleolus),
                tracer = c(tracer)) %>%
    filter(object != 0) %>%
    rowwise() %>%
    mutate(nucleolus_internal = min(nucleolus_internal, 254),
           nucleolus_external = min(nucleolus_external, 254)) %>%
    ungroup() %>%
    group_by(object, nucleolus_internal, nucleolus_external) %>%
    summarise(n = n(),
              tracer = mean(tracer),
              nucleolus = mean(nucleolus)) %>%
    ungroup() %>%
    group_by(object) %>%
    mutate(tracer_enrichment = log2(tracer / weighted.mean(tracer, n)),
           nucleolus_enrichment = log2(nucleolus / weighted.mean(nucleolus, n))) %>%
    ungroup() %>%
    filter(n > n_min) %>%
    mutate(distance = nucleolus_external - nucleolus_internal) %>%
    arrange(object, distance) %>%
    add_column(sample = metadata$sample[i],
               replicate = metadata$replicate[i],
               cell = metadata$cell[i],
               condition = metadata$condition[i],
               target = metadata$target[i],
               slide = metadata$slide[i]) %>%
    mutate(cell_id = paste(sample, object, sep = "_"))
  
  tib
  
}

tib_intensities <- lapply(1:nrow(metadata), LoadPGMMaps) %>%
  purrr::reduce(bind_rows)

# Filter out cells that did not pass the cell size filter
tib_intensities <- tib_intensities %>%
  dplyr::filter(cell_id %in% (tib_cells %>% 
                                filter(size_filter == T) %>%
                                pull(cell_id)))

```

Plots.

```{r plot intensities, fig.width = 10, fig.height = 7}

# Filter very bad cells
# - only positive / negative scores
# - extreme values
remove_samples <- tib_intensities %>% 
  gather(key, value, nucleolus_enrichment, tracer_enrichment) %>%
  mutate(key = str_remove(key, "_enrichment")) %>%
  filter(distance > -11 & distance <= 11) %>% 
  group_by(key, target, cell, cell_id) %>% 
  dplyr::summarise(q1 = quantile(value, 0.1), 
                   q5 = quantile(value, 0.5), 
                   q9 = quantile(value, 0.9)) %>%
  filter(q1 > 0.5 | q9 < 0 | q1 < -2.5 | q9 > 3) %>%
  pull(cell_id)


# Plot intensity per cell
tib_gather <- tib_intensities %>%
  filter(! cell_id %in% remove_samples) %>%  
  gather(key, value, nucleolus, tracer) %>%
  filter(distance > -14 & distance <= 14) %>% 
  mutate(key = factor(key, levels = c("tracer", "nucleolus")))

tib_gather %>%  
  ggplot(aes(x = distance, y = value, col = key, 
             group = interaction(key, cell_id))) +
  geom_hline(yintercept = 0, col = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, col = "black", linetype = "dashed") +
  geom_line() +
  facet_grid(target ~ condition, scales = "free_y") +
  xlab("Distance to nucleolus (pixel)") +
  ylab("Intensity (A.U.)") +
  coord_cartesian(xlim = c(-10, 10)) +
  scale_color_grey() +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Plot intensity per cell - for enrichments
tib_gather <- tib_intensities %>%
  filter(! cell_id %in% remove_samples) %>%  
  gather(key, value, nucleolus_enrichment, tracer_enrichment) %>%
  filter(distance > -14 & distance <= 14) %>% 
  mutate(key = str_remove(key, "_enrichment")) %>%
  mutate(key = factor(key, levels = c("tracer", "nucleolus")))

# Number of cells
tib_gather %>%
  group_by(cell, condition, target, cell_id) %>%
  dplyr::summarise(data_points = n()) %>%
  group_by(cell, condition, target) %>%
  dplyr::summarise(n = n(),
                   data_points = sum(data_points)) %>%
  print(n = 10)

tib_gather %>%  
  ggplot(aes(x = distance, y = value, col = key, 
             group = interaction(key, cell_id))) +
  geom_hline(yintercept = 0, col = "black") +
  geom_vline(xintercept = 0, col = "black") +
  geom_line(alpha = 0.2) +
  stat_summary(fun = mean, aes(group = key), 
               geom = "line", size = 2) +
  facet_grid(target ~ condition, scales = "free_y") +
  xlab("Distance to nucleolus (pixel)") +
  ylab("Enrichment over mean (log2)") +
  coord_cartesian(xlim = c(-10, 10),
                  ylim = c(-2.1, 2.3)) +
  #scale_color_grey() +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Means only
tib_gather %>%  
  #filter(! condition %in% c("t0h", "t21h")) %>%
  filter(! condition %in% c("t0h")) %>%
  ggplot(aes(x = distance, y = value, col = condition, linetype = target,
             fill = condition, group = interaction(target, condition))) +
  geom_hline(yintercept = 0, col = "black") +
  geom_vline(xintercept = 0, col = "black") +
  stat_summary(fun = mean, geom = "line", size = 1) +
  # stat_summary(fun.data = mean_se, geom = "errorbar",
  #              col = "darkgrey", linetype = "solid") +
  stat_summary(fun.data = mean_se, geom = "ribbon",
               col = NA, alpha = 0.2) +
  facet_grid(key ~ .) +
  xlab("Distance to nucleolus (pixel)") +
  ylab("Enrichment over mean (log2)") +
  coord_cartesian(xlim = c(-10, 10),
                  ylim = c(-1.1, 2.1)) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  #scale_color_grey() +
  scale_color_manual(values = c("black", "red1", "red2", "red3", "red4")) +
  scale_fill_manual(values = c("black", "red1", "red2", "red3", "red4")) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Change color and linetype
tib_gather %>%  
  #filter(! condition %in% c("t0h", "t21h")) %>%
  filter(! condition %in% c("t0h")) %>%
  ggplot(aes(x = distance, y = value, col = target, linetype = condition,
             fill = target, group = interaction(target, condition))) +
  geom_hline(yintercept = 0, col = "black") +
  geom_vline(xintercept = 0, col = "black") +
  stat_summary(fun = mean, geom = "line", size = 1) +
  # stat_summary(fun.data = mean_se, geom = "errorbar",
  #              col = "darkgrey", linetype = "solid") +
  stat_summary(fun.data = mean_se, geom = "ribbon",
               col = NA, alpha = 0.2) +
  facet_grid(key ~ .) +
  xlab("Distance to nucleolus (pixel)") +
  ylab("Enrichment over mean (log2)") +
  coord_cartesian(xlim = c(-10, 10),
                  ylim = c(-1.1, 2.1)) +
  scale_linetype_manual(values = c("solid", "longdash", "dashed", "dotdash", "dotted")) +
  #scale_color_grey() +
  scale_color_manual(values = c("black", "red")) +
  scale_fill_manual(values = c("black", "red")) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


### Conclusion

Several observations:

  * Ki67 m6A-Tracer is enriched near Ki67 protein. That's good.
  * This enrichment is almost equal between HCT116 and K562 cells.
  
Overall, this is a positive result.


### SessionInfo

```{r sessionInfo}

sessionInfo()

```














