---
title: "4DN DamID - Ki67 project - Replication timing"
author: "Tom van Schaik"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

### Project

Mapping of Ki67 interactions with the genome and comparison with lamina 
interactions.


### Introduction

Processing of replication timing. I already trimmed, mapped and counted the 
reads in 50kb bins. In this document, I will do the remaining processing:

  * Counts-per-million normalization
  * Log2-ratios of early / late
  * Quantile normalization
  * Optionally: smoothing
  
I will write the data as .rds file, and prepare bigwig files.


### Method

See introduction.


### Set-up

Set the parameters and list the data.

```{r set-up, warning = F, message = F, cache = T}

# Load dependencies - without warnings / messages
library(tidyverse)
library(GenomicRanges)
library(rtracklayer)
library(ggplot2)
library(preprocessCore) # quantile normalization

# Prepare output 
output_dir <- "ts210802_further_processing_replication_timing"
dir.create(output_dir, showWarnings = FALSE)


# Load input
chromosomes <- c(paste0("chr", 1:22), "chrX")


# Prepare seqnames
chrom_sizes <- read_tsv("~/mydata/data/genomes/GRCh38/hg38.chrom.sizes",
                        col_names = c("seqnames", "length"))

```

```{r knits setup}
library(knitr)
opts_chunk$set(fig.width = 5, fig.height = 3.5, cache = T,
               dev=c('png', 'pdf'), 
               fig.path = file.path(output_dir, "figures/")) 
pdf.options(useDingbats = FALSE)
```

```{r functions}


```


### 1. Load counts

Load counts.

```{r load counts}

# Read counts
counts <- read_tsv("ts210802_repliseq/deeptools/deeptools_bam_coverage.tab")

# Reformatting
counts <- counts %>%
  rename_at(1:3, ~ c("seqnames", "start", "end")) %>%
  rename_at(4:ncol(.), str_remove_all, "'") %>%
  mutate(start = start + 1) %>%
  filter(seqnames %in% chromosomes) %>%
  mutate(seqnames = factor(seqnames, levels = chromosomes)) %>%
  arrange(seqnames, start) %>%
  mutate(seqnames = as.character(seqnames))

```


### 2. Normalization

First, normalize to cpm. Also, bins < 10 reads are removed from further
analyses.

```{r normalize to cpm}

# Prepare metadata
samples <- names(counts)[4:ncol(counts)]

metadata <- tibble(sample = samples) %>%
  mutate(fraction = ifelse(str_detect(sample, "_e_"),
                           "early", "late"),
         iaa = ifelse(str_detect(sample, "no_iaa"),
                      "no_iaa", "iaa"),
         lane = ifelse(str_detect(sample, "L001"),
                       "L001", "L002"),
         rep = ifelse(str_detect(sample, "r1"),
                      "r1", "r2")) %>%
  mutate(experiment = paste("HCT116", iaa, rep, sep = "_"),
         experiment_fraction = paste("HCT116", iaa, fraction, rep, sep = "_"))

# Combine lanes
counts_lanes <- counts %>%
  mutate(
    HCT116_no_iaa_early_r1 = 
      hct116_no_iaa_r1_e_S35_L001_R1 + 
      hct116_no_iaa_r1_e_S35_L002_R1,
    HCT116_no_iaa_late_r1 = 
      hct116_no_iaa_r1_l_S36_L001_R1 + 
      hct116_no_iaa_r1_l_S36_L002_R1,
    HCT116_iaa_early_r1 = 
      hct116_iaa_r1_e_S37_L001_R1 + 
      hct116_iaa_r1_e_S37_L002_R1,
    HCT116_iaa_late_r1 = 
      hct116_iaa_r1_l_S38_L001_R1 + 
      hct116_iaa_r1_l_S38_L002_R1,
    HCT116_no_iaa_early_r2 = 
      hct116_no_iaa_r2_e_S43_L001_R1 + 
      hct116_no_iaa_r2_e_S43_L002_R1,
    HCT116_no_iaa_late_r2 = 
      hct116_no_iaa_r2_l_S44_L001_R1 + 
      hct116_no_iaa_r2_l_S44_L002_R1,
    HCT116_iaa_early_r2 = 
      hct116_iaa_r2_e_S45_L001_R1 + 
      hct116_iaa_r2_e_S45_L002_R1,
    HCT116_iaa_late_r2 = 
      hct116_iaa_r2_l_S46_L001_R1 + 
      hct116_iaa_r2_l_S46_L002_R1
  ) %>%
  dplyr::select(-contains("_S"))


# Normalize to cpm
cpm <- function(x, min_count = 0) {
  # Normalize vector to cpm (total count: 1M)
  x_cpm <- x / sum(x) * 1e6
  x_cpm[x < min_count] <- NA
  
  x_cpm
}

counts_norm <- counts_lanes %>%
  mutate_at(all_of(unique(metadata$experiment_fraction)), cpm)


# Finally, remove values for rows with too few reads (in absolute numbers)
min_reads <- 10

idx_not_enough_reads <- counts_lanes %>%
  dplyr::select(4:ncol(.)) %>%
  mutate_all(function(x) x >= min_reads) %>%
  mutate(sum = rowSums(.),
         idx = sum <= 3) %>%
  pull(idx)

counts_norm[idx_not_enough_reads, 4:ncol(counts_norm)] <- NA

  
```

Then, calculate log2-ratio of E/L.

```{r early late ratio}

# Log2-ratio
norm <- counts_norm %>%
  mutate(
    HCT116_no_iaa_early_r1 = 
      log2((HCT116_no_iaa_early_r1 + 1) / 
             (HCT116_no_iaa_late_r1 + 1)),
    HCT116_iaa_early_r1 = 
      log2((HCT116_iaa_early_r1 + 1) / 
             (HCT116_iaa_late_r1 + 1)),
    HCT116_no_iaa_early_r2 = 
      log2((HCT116_no_iaa_early_r2 + 1) / 
             (HCT116_no_iaa_late_r2 + 1)),
    HCT116_iaa_early_r2 = 
      log2((HCT116_iaa_early_r2 + 1) / 
             (HCT116_iaa_late_r2 + 1))) %>%
  dplyr::select(-contains("late"))

# Quantile normalization
norm_quant <- norm
norm_quant[, 4:ncol(norm_quant)] <- norm_quant %>%
  dplyr::select(4:ncol(.)) %>%
  as.matrix() %>%
  normalize.quantiles(.)
  

```

Finally, get the mean between the replicates to use in downstream analyses.

```{r mean between replicates}

# Average between replicates
norm_average <- norm_quant %>%
  rowwise() %>%
  mutate(
    HCT116_no_iaa = 
      mean(c(HCT116_no_iaa_early_r1, 
             HCT116_no_iaa_early_r2), 
           na.rm = T),
    HCT116_iaa = 
      mean(c(HCT116_iaa_early_r1, 
             HCT116_iaa_early_r2), 
           na.rm = T)
  ) %>%
  ungroup() %>%
  dplyr::select(-contains("_r"))

```

### 3. Create bigwigs

Finally, create bigwigs of the (quantile) normalized tracks.

```{r prepare bigwigs}

# Function
ExportBWFromTibble <- function(tib, bigwig_dir, column = "diff", name = NULL) {
  
  # Set name as column
  if (is.null(name)) name <- column
  
  # Get GRanges from Tibble
  gr <- tib %>%
    dplyr::select(1:3, all_of(column)) %>%
    drop_na() %>%
    dplyr::rename_at(4, ~ "score") %>%
    as(., "GRanges")
  
  # Exception: LADs and SPIN states
  if (grepl("LAD", column)) {
    gr$score <- ifelse(gr$score == "LAD", 1, 0)
  }
  
  if (grepl("SPIN", column)) {
    return(NULL)
  }
  
  # Add chromosome information
  seqlengths(gr) <- chrom_sizes$length[match(seqlevels(gr),
                                             chrom_sizes$seqnames)]
  
  # Write bigwig
  export.bw(object = gr,
            con = file.path(bigwig_dir, paste0(name, ".bw")))
  
}

BWsFromTibble <- function(tib) {
  invisible(lapply(names(tib)[4:ncol(tib)],
                   function(sample) {
                     ExportBWFromTibble(tib,
                                        bigwig_dir = bigwig_dir,
                                        column = sample)
                   }
  ))
}

# Prepare bigwigs - normalized
bigwig_dir <- file.path(output_dir, "bigwig")
dir.create(bigwig_dir, showWarnings = F)

BWsFromTibble(norm_quant)
BWsFromTibble(norm_average)

# Prepare bigwigs - counts
bigwig_dir <- file.path(output_dir, "bigwig_counts")
dir.create(bigwig_dir, showWarnings = F)

BWsFromTibble(counts_norm)

```


### 4. Plots

I want to make some plots.

```{r plots}

# Correlation between replicates - before and after quantile
norm %>%
  ggplot(aes(x = HCT116_no_iaa_early_r1, y =HCT116_no_iaa_early_r2)) +
  geom_point(alpha = 0.1) +
  geom_abline(intercept = 0, slope = 1, col = "red") +
  ggtitle("Before quantile") +
  theme_bw() +
  theme(aspect.ratio = 1)

norm_quant %>%
  ggplot(aes(x = HCT116_no_iaa_early_r1, y =HCT116_no_iaa_early_r2)) +
  geom_point(alpha = 0.1) +
  geom_abline(intercept = 0, slope = 1, col = "red") +
  ggtitle("After quantile") +
  theme_bw() +
  theme(aspect.ratio = 1)


# Correlation between +IAA and -IAA - before and after quantile
norm %>%
  ggplot(aes(x = HCT116_no_iaa_early_r1, y =HCT116_iaa_early_r2)) +
  geom_point(alpha = 0.1) +
  geom_abline(intercept = 0, slope = 1, col = "red") +
  ggtitle("Before quantile") +
  theme_bw() +
  theme(aspect.ratio = 1)

norm_quant %>%
  ggplot(aes(x = HCT116_no_iaa_early_r1, y =HCT116_iaa_early_r2)) +
  geom_point(alpha = 0.1) +
  geom_abline(intercept = 0, slope = 1, col = "red") +
  ggtitle("After quantile") +
  theme_bw() +
  theme(aspect.ratio = 1)

norm_average %>%
  ggplot(aes(x = HCT116_no_iaa, y =HCT116_iaa)) +
  geom_point(alpha = 0.1) +
  geom_abline(intercept = 0, slope = 1, col = "red") +
  ggtitle("Average after quantile") +
  theme_bw() +
  theme(aspect.ratio = 1)

```

These plots show that some normalization is good. Quantile might be a bit
overkill, but this is always used for replication timing as I understood.


### Conclusions

I previously received repliseq data from Ethan. However, there was some
confusion whether his analyses were sound. I decided to repeat everything to
be sure of the results. This analysis is mostly using the same strategy / 
tools as the pA-DamID processing. 

Regarding the previous conclusions ("nothing really changes except near 
centromeres"), I can see that this conclusion is not affected. Furthermore,
I did solve the issues: strange coverage at centromeres + no coverage at X
chromosome.


### Session info

```{r save data}

# Save RDS files of the ranges
saveRDS(norm_quant, file.path(output_dir, "tib_norm_quant.rds"))
saveRDS(norm_average, file.path(output_dir, "tib_norm_average.rds"))
saveRDS(metadata, file.path(output_dir, "tib_repliseq_metadata.rds"))

```

```{r session info}
sessionInfo()
```

