---
title: "Intesity of Ki-67 around centromeres"
author: "Tom van Schaik"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---


### Introduction

The goal of this analysis is to determine the enrichment of Ki-67 around 
centromeres. I have generated data with and without actinomycin D treatment.


### Setup

Load the required libraries.

```{r setup, message = F, warning = F}

# Load libraries
library(ggplot2)
library(tidyverse)
library(ggbeeswarm)
library(pixmap)
library(raster)

# Input data
results_dir <- c("../ts220405_E1936_confocal/ts220405_RPE_ActD/ts220405_RPE_ActD_analysis/",
                 "../ts220414_E1936_confocal/ts220414_RPE_ActD/ts220414_RPE_ActD_analysis/")

# Prepare output directory
output_dir <- "ts220412_ki67_at_centromeres"
dir.create(output_dir, showWarnings = FALSE)

# Prepare output files

```

```{r knits setup}
library(knitr)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(output_dir, "figures/"),
               cache = T)
pdf.options(useDingbats = FALSE)
```

```{r functions}

ReadPGN <- function(file.name) {
  
  # Get image, only grey values
  # suppressWarnings(read.pnm(file.name)@grey*255)
  
  values(raster(file.name))
  
}

```


### Strategy

...


### 1) Prepare metadata

List the different samples.

```{r prepare metadata}

# List and organize all files
metadata <- tibble(cell_statistics = dir(results_dir, recursive = T, 
                                         full.names = T,
                                         pattern = "cells_statistics")) %>%
  filter(! grepl("MKI67IP", cell_statistics)) %>%
  mutate(sample = str_remove(basename(cell_statistics),
                             "_cells_.*")) %>%
  separate(sample, remove = F,
           c("cell", "antibody1", "antibody2",
             "condition", "replicate", "slide")) %>%
  mutate(replicate = factor(replicate, levels = c("r1", "r2", "r3")),
         cell = factor("RPE", levels = c("RPE")),
         condition = factor(condition, levels = c("DMSO", "ActD")),
         slide = ifelse(grepl("ts220414", cell_statistics),
                        paste0(slide, "_220414"),
                        paste0(slide, "_220405")),
         sample = paste(cell, condition, replicate, slide, sep = "_")) %>%
  mutate(objects = str_replace(cell_statistics, 
                               "cells_statistics.csv",
                               "dapi_segment.tiff"),
         centromeres_external = str_replace(objects, 
                                            "dapi_segment", 
                                            "centromeres_external"),
         centromeres = str_replace(objects, 
                                   "dapi_segment", 
                                   "centromeres_smooth"),
         ki67 = str_replace(objects, 
                            "dapi_segment", 
                            "Ki67_smooth"),
         centromeres_objects = str_replace(objects, 
                                           "dapi_segment", 
                                           "centromeres_objects")) %>%
  arrange(cell, condition, replicate, slide)

# Print metadata
metadata %>% 
  print(n = Inf)

```


### 2) Load cell sizes and create a cell table

Load the cells and show the sizes. Possibly filter on those.

```{r load cells, message = F, warning = F}

# Load all cells
LoadCells <- function(i) {
  tib <- read_csv(metadata$cell_statistics[i]) %>%
    dplyr::select(c(1:2)) %>%
    rename_all(~c("volume", "surface")) %>%
    add_column(object = 1:nrow(.),
               sample = metadata$sample[i],
               replicate = metadata$replicate[i],
               condition = metadata$condition[i],
               cell = metadata$cell[i], 
               slide = metadata$slide[i]) %>%
    mutate(cell_id = paste(sample, object, sep = "_"))
  
  tib
  
}

tib_cells <- lapply(1:nrow(metadata), LoadCells) %>%
  purrr::reduce(bind_rows)

```

Plot and filter these cells.

```{r plot cells, fig.width = 5, fig.height = 3.5}

# Set size limits
limits_size <- c(25, 150)

# Plot cell volume
tib_cells %>%
  ggplot(aes(x = condition, y = volume, col = replicate)) +
  geom_hline(yintercept = 0, alpha = 0) +
  geom_hline(yintercept = limits_size, col = "black", linetype = "dashed") +
  geom_quasirandom() +
  geom_boxplot(fill = NA, col = "black", outlier.shape = NA) +
  facet_grid(. ~ condition, scales = "free", space = "free") +
  xlab("") +
  ylab("Volume slice (micron)") +
  theme_bw() +
  theme(#aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Filter cells for cutoff 
tib_cells <- tib_cells %>%
  mutate(size_filter = volume >= limits_size[1] & 
           volume <= limits_size[2])

tib_cells %>%
  filter(size_filter == T) %>%
  ggplot(aes(x = condition, y = volume, col = replicate)) +
  geom_hline(yintercept = 0, alpha = 0) +
  geom_hline(yintercept = limits_size, col = "black", linetype = "dashed") +
  geom_quasirandom() +
  geom_boxplot(fill = NA, col = "black", outlier.shape = NA) +
  facet_grid(. ~ condition, scales = "free", space = "free") +
  xlab("") +
  ylab("Volume slice (micron)") +
  theme_bw() +
  theme(#aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


### 3) Load the pgm maps

Load the .pgm files for the various channels.

```{r load pgm, message = F, warning = F}

# Load all cells
LoadPGMMaps <- function(i, n_min = 5) {
  
  # Load objects
  objects <- ReadPGN(metadata$objects[i])
  centromeres <- ReadPGN(metadata$centromeres[i])
  centromeres_external <- ReadPGN(metadata$centromeres_external[i])
  ki67 <- ReadPGN(metadata$ki67[i])
  centromeres_objects <- ReadPGN(metadata$centromeres_objects[i])
  
  tib <- tibble(object = c(objects),
                centromeres_external = c(centromeres_external),
                centromeres = c(centromeres),
                ki67 = c(ki67),
                centromeres_objects = c(centromeres_objects)) %>%
    filter(object != 0) %>%
    rowwise() %>%
    mutate(centromeres_external = min(centromeres_external, 255)) %>%
    ungroup() %>%
    group_by(object, centromeres_external) %>%
    summarise(n = n(),
              ki67 = mean(ki67), 
              centromeres = mean(centromeres)) %>%
    ungroup() %>%
    group_by(object) %>%
    mutate(ki67_enrichment = log2(ki67 / weighted.mean(ki67, n)),
           centromeres_enrichment = log2(centromeres / weighted.mean(centromeres, n))) %>%
    ungroup() %>%
    filter(n > n_min) %>%
    mutate(distance = centromeres_external) %>%
    arrange(object, distance) %>%
    add_column(sample = metadata$sample[i],
               replicate = metadata$replicate[i],
               cell = metadata$cell[i],
               condition = metadata$condition[i],
               slide = metadata$slide[i]) %>%
    mutate(cell_id = paste(sample, object, sep = "_"))
  
  tib
  
}

tib_intensities <- lapply(1:nrow(metadata), LoadPGMMaps) %>%
  purrr::reduce(bind_rows)

# Filter out cells that did not pass the cell size filter
tib_intensities <- tib_intensities %>%
  dplyr::filter(cell_id %in% (tib_cells %>% 
                                filter(size_filter == T) %>%
                                pull(cell_id)))

```

Plots.

```{r plot intensities, fig.width = 10, fig.height = 7}

# Filter very bad cells
# - no Ki-67 signal
# - only positive / negative scores
# - extreme values
remove_samples_1 <- tib_intensities %>% 
  gather(key, value, centromeres_enrichment, ki67_enrichment) %>%
  mutate(key = str_remove(key, "_enrichment")) %>%
  filter(distance > -11 & distance <= 11) %>% 
  group_by(key, cell, cell_id) %>% 
  dplyr::summarise(q1 = quantile(value, 0.1), 
                   q5 = quantile(value, 0.5), 
                   q9 = quantile(value, 0.9)) %>%
  filter(q1 > 0.5 | q9 < 0 | q1 < -100 | q9 > 100) %>%
  pull(cell_id)
 
remove_samples_2 <- tib_intensities %>% 
  group_by(sample, cell, cell_id) %>% 
  dplyr::summarise(ki67 = mean(ki67)) %>%
  filter(ki67 < 20) %>%
  pull(cell_id)

remove_samples <- unique(c(remove_samples_1,
                           remove_samples_2))


# Plot intensity per cell
tib_gather <- tib_intensities %>%
  filter(! cell_id %in% remove_samples) %>%  
  gather(key, value, centromeres, ki67) %>%
  filter(distance >= 0 & distance <= 16) %>% 
  mutate(key = factor(key, levels = c("centromeres", "ki67")))

tib_gather %>%  
  ggplot(aes(x = distance, y = value, col = key, 
             group = interaction(key, cell_id))) +
  geom_hline(yintercept = 0, col = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, col = "black", linetype = "dashed") +
  geom_line() +
  facet_grid(key ~ condition, scales = "free_y") +
  xlab("Distance to centromeres (pixel)") +
  ylab("Intensity (A.U.)") +
  coord_cartesian(xlim = c(0, 15)) +
  scale_color_grey() +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


# Plot intensity per cell - for enrichments
tib_gather <- tib_intensities %>%
  filter(! cell_id %in% remove_samples) %>%  
  gather(key, value, centromeres_enrichment, ki67_enrichment) %>%
  filter(distance >= 0 & distance <= 16) %>% 
  mutate(key = str_remove(key, "_enrichment")) %>%
  mutate(key = factor(key, levels = c("centromeres", "ki67")))

# Number of cells
tib_gather %>%
  group_by(cell, condition, cell_id) %>%
  dplyr::summarise(data_points = n()) %>%
  group_by(cell, condition) %>%
  dplyr::summarise(n = n(),
                   data_points = sum(data_points)) %>%
  print(n = 10)

tib_gather %>%  
  ggplot(aes(x = distance, y = value,
             group = interaction(key, cell_id))) +
  geom_hline(yintercept = 0, col = "black") +
  geom_vline(xintercept = 0, col = "black") +
  geom_line(alpha = 0.2) +
  stat_summary(fun = mean, aes(group = key), 
               geom = "line", size = 2) +
  facet_grid(key ~ condition, scales = "free_y") +
  xlab("Distance to centromeres (pixel)") +
  ylab("Enrichment over mean (log2)") +
  coord_cartesian(xlim = c(0, 15)) +
  #scale_color_grey() +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Means only
tib_gather %>%  
  ggplot(aes(x = distance, y = value, col = condition, linetype = key,
             fill = condition, group = interaction(key, condition))) +
  geom_hline(yintercept = 0, col = "black") +
  geom_vline(xintercept = 0, col = "black") +
  stat_summary(fun = mean, geom = "line", size = 1) +
  # stat_summary(fun.data = mean_se, geom = "errorbar",
  #              col = "darkgrey", linetype = "solid") +
  stat_summary(fun.data = mean_se, geom = "ribbon",
               col = NA, alpha = 0.2) +
  facet_grid(key ~ ., scales = "free_y") +
  xlab("Distance to centromeres (pixel)") +
  ylab("Enrichment over mean (log2)") +
  coord_cartesian(xlim = c(0, 15)) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  #scale_color_grey() +
  scale_color_manual(values = c("black", "red1", "red2", "red3", "red4")) +
  scale_fill_manual(values = c("black", "red1", "red2", "red3", "red4")) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


### 4) Enrichment at individual centromeres

The figures above show the average Ki-67 at all centromeres per cell. I also
want to determine the enrichment at individual centromeres. This will just be
one number rather than a distance function.

```{r load centromere data}

# Determine enrichment at centromeres
# Load all cells
LoadCentromeres <- function(i) {
  
  # Load objects
  objects <- ReadPGN(metadata$objects[i])
  centromeres <- ReadPGN(metadata$centromeres[i])
  centromeres_external <- ReadPGN(metadata$centromeres_external[i])
  ki67 <- ReadPGN(metadata$ki67[i])
  centromeres_objects <- ReadPGN(metadata$centromeres_objects[i])
  
  tib <- tibble(object = c(objects),
                centromeres_external = c(centromeres_external),
                centromeres = c(centromeres),
                ki67 = c(ki67),
                centromeres_objects = c(centromeres_objects)) %>%
    filter(object != 0) %>%
    group_by(object, centromeres_objects) %>%
    summarise(n = n(),
              ki67 = mean(ki67), 
              centromeres = mean(centromeres)) %>%
    ungroup() %>%
    group_by(object) %>%
    mutate(ki67_mean = weighted.mean(ki67, n),
           ki67_enrichment = log2(ki67 / weighted.mean(ki67, n)),
           centromeres_enrichment = log2(centromeres / weighted.mean(centromeres, n))) %>%
    ungroup() %>%
    filter(centromeres_objects != 0) %>%
    arrange(object, ki67_enrichment) %>%
    add_column(sample = metadata$sample[i],
               replicate = metadata$replicate[i],
               cell = metadata$cell[i],
               condition = metadata$condition[i],
               slide = metadata$slide[i]) %>%
    mutate(cell_id = paste(sample, object, sep = "_"))
  
  tib
  
}

tib_centromeres <- lapply(1:nrow(metadata), LoadCentromeres) %>%
  purrr::reduce(bind_rows)

# Filter out cells that did not pass the cell size filter
tib_centromeres <- tib_centromeres %>%
  dplyr::filter(cell_id %in% (tib_cells %>% 
                                filter(size_filter == T) %>%
                                pull(cell_id)))

```

Next, create some plots

```{r plot centromere data, fig.width = 5, fig.height = 3.5}

# Plot all data
tib_centromeres %>%
  filter(! cell_id %in% remove_samples) %>%  
  ggplot(aes(x = condition, y = ki67_enrichment, col = condition)) +
  geom_hline(yintercept = 0, col = "black") +
  geom_quasirandom() +
  geom_boxplot(outlier.shape = NA, col = "black", fill = NA) +
  scale_color_manual(values = c("grey50", "red"), guide = "none") +
  xlab("") +
  ylab("Ki-67 enrichment over mean (log2)") +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


### Conclusion

Several observations:

  * Ki-67 is clearly enriched at (most) centromeres.
  * This enrichment extends beyond the CENPA signal.
  
Overall, this is a positive result.


### SessionInfo

```{r sessionInfo}

sessionInfo()

```














