---
title: "Intesity of Ki67 after auxin addition"
author: "Tom van Schaik"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---


### Introduction

The goal of this analysis is to determine the downregulation of Ki67 after 
auxin-mediated knockdown. I have images of HCT116 Ki67 AID 2-4 (Takagi, et al.). 
Microscopy looked good for the GFP fusion, and less so for the antibody 
staining. Let's quantify.

Update ts210408: I repeated microscopy of Ki67-AID IAA-mediated depletion, this
time with two different antibodies. However, I took 2D images rather than 3D 
images. Let's quantify. I might want to make 3D images in the end, but on the
other hand this could be enough.


### Setup

Load the required libraries.

```{r setup, message = F, warning = F}

# Load libraries
library(ggplot2)
library(tidyverse)
library(ggbeeswarm)

# Input data
results_dir <- c("../ts210402_E1546_confocal_Ki67_antibodies/ts210402_HCT116_Ki67AID_Ki67_depletion/ts210402_HCT116_Ki67AID_Ki67_depletion_analysis/")

# Prepare output directory
output_dir <- "ts210609_ki67_intensity_following_auxin_depletion_other_antibodies_2D"
dir.create(output_dir, showWarnings = FALSE)

# Prepare output files

```

```{r knits setup}
library(knitr)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(output_dir, "figures/"),
               cache = T)
pdf.options(useDingbats = FALSE)
```

```{r functions}


```


### Strategy

...


### 1) Prepare metadata

List the different samples.

```{r prepare metadata}

# List and organize all files
metadata <- tibble(cell_statistics = dir(results_dir, recursive = T, 
                                         full.names = T,
                                         pattern = "dapi_statistics")) %>%
  mutate(sample = str_remove(basename(cell_statistics),
                             "_dapi.*"),
         sample = str_remove(sample, "M_")) %>%
  separate(sample, remove = F,
           c("cell", "replicate", "target", "condition", "slide")) %>%
  mutate(condition = recode(condition,
                            minusIAA = "ctrl",
                            plusIAA = "IAA"),
         condition = factor(condition, levels = c("ctrl", "IAA")),
         replicate = factor(replicate, levels = c("r1", "r2", "r3")),
         target = factor(target, levels = c("HPA", "NOV"))) %>%
  mutate(dapi = str_replace(cell_statistics,
                            "M_", "Q_"),
         dapi = str_remove(dapi, "_statistics"),
         gfp = str_replace(dapi, "dapi", "mki67ip"),
         ki67 = str_replace(dapi, "dapi", "ki67")) %>%
  arrange(target, condition, replicate, slide)

# Print metadata
metadata %>% 
  print(n = Inf)

```


### 2) Load cell sizes and create a cell table

Load the cells and show the sizes. Possibly filter on those.

```{r load cells, message = F, warning = F}

# Load all cells
LoadCells <- function(i) {
  tib <- read_tsv(metadata$cell_statistics[i]) %>%
    dplyr::select(c(2, 6, 8)) %>%
    rename_all(~c("object", "volume", "surface")) %>%
    add_column(sample = metadata$sample[i],
               condition = metadata$condition[i],
               slide = metadata$slide[i],
               target = metadata$target[i],
               replicate = metadata$replicate[i]) %>%
    mutate(cell_id = paste(sample, object, sep = "_"))
  
  tib
  
}

tib_cells <- lapply(1:nrow(metadata), LoadCells) %>%
  purrr::reduce(bind_rows)

```

Plot and filter these cells.

```{r plot cells, fig.width = 8, fig.height = 3.5}

# Set size limits
limits_size <- c(8e3, 3e4)

# Plot cell volume
tib_cells %>%
  ggplot(aes(x = condition, y = volume, col = replicate)) +
  geom_hline(yintercept = limits_size, col = "black", linetype = "dashed") +
  geom_quasirandom() +
  geom_boxplot(fill = NA, col = "black", outlier.shape = NA) +
  xlab("") +
  ylab("Volume (pixels)") +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Filter cells for cutoff
tib_cells <- tib_cells %>%
  mutate(size_filter = volume >= limits_size[1] & 
           volume <= limits_size[2])

tib_cells %>%
  filter(size_filter == T) %>%
  ggplot(aes(x = condition, y = volume, col = replicate)) +
  geom_hline(yintercept = limits_size, col = "black", linetype = "dashed") +
  geom_quasirandom() +
  geom_boxplot(fill = NA, col = "black", outlier.shape = NA) +
  xlab("") +
  ylab("Volume (pixels)") +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


### 3) Load the intensities

Load the intensities of the various channels (DAPI, gfp, Ki67).

```{r load intensities, message = F, warning = F}

# Load all cells
LoadIntensities <- function(i) {
  
  # Load sizes
  tib <- read_tsv(metadata$dapi[i]) %>%
    dplyr::select(2, 6) %>%
    rename_all(~c("object", "dapi")) %>%
    add_column(sample = metadata$sample[i],
               condition = metadata$condition[i],
               target = metadata$target[i],
               slide = metadata$slide[i],
               replicate = metadata$replicate[i]) %>%
    mutate(cell_id = paste(sample, object, sep = "_"))
  
  # Load overlap with cells and lamina
  tib_gfp <- read_tsv(metadata$gfp[i]) %>%
    dplyr::select(6) %>%
    rename_all(~c("gfp"))
  
  tib_ki67 <- read_tsv(metadata$ki67[i]) %>%
    dplyr::select(6) %>%
    rename_all(~c("ki67"))
    
  # Combine all
  tib <- bind_cols(tib, tib_gfp, tib_ki67) %>%
    arrange(condition, slide, object)
  
  tib
  
}

tib_intensities <- lapply(1:nrow(metadata), LoadIntensities) %>%
  purrr::reduce(bind_rows)

# Filter out cells that did not pass the cell size filter
tib_intensities <- tib_intensities %>%
  dplyr::filter(cell_id %in% (tib_cells %>% 
                                filter(size_filter == T) %>%
                                pull(cell_id)))

```

Plot various numbers of the targets.

```{r plot targets, fig.width = 7, fig.height = 7}

# Plot intensity per cell
value_norm_limit <- 5

tib_gather <- tib_intensities %>%  
  mutate(volume = tib_cells$volume[match(tib_intensities$cell_id,
                                         tib_cells$cell_id)]) %>%
  filter(cell_id != "minusIAA_007_34") %>%
  gather(key, value, dapi, gfp, ki67) %>%
  mutate(key = factor(key, levels = c("dapi", "gfp", "ki67")),
         value_norm = value * volume) %>%
  group_by(key, target, replicate) %>%
  mutate(value_norm = value_norm / quantile(value_norm, 0.8) * 10,
         shape = ifelse(value_norm < value_norm_limit, 
                        "-", "cut"),
         value_norm_cut = ifelse(value_norm < value_norm_limit, 
                                 value_norm, 0.5)) %>%
  ungroup()

tib_gather %>%  
  ggplot(aes(x = condition, y = value, col = replicate)) +
  geom_hline(yintercept = 0, col = "black", linetype = "solid") +
  geom_quasirandom() +
  geom_boxplot(fill = NA, col = "black", outlier.shape = NA) +
  facet_grid(key ~ target, scales = "free_y") +
  xlab("") +
  ylab("Intensity (A.U.)") +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Also, normalize by cell volume
tib_gather %>%  
  ggplot(aes(x = condition, y = value_norm, col = replicate)) +
  geom_hline(yintercept = 0, col = "black", linetype = "solid") +
  geom_quasirandom() +
  geom_boxplot(fill = NA, col = "black", outlier.shape = NA) +
  facet_grid(key ~ target, scales = "free_y") +
  xlab("") +
  ylab("Intensity (A.U.)") +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

tib_gather %>%  
  ggplot(aes(x = condition, y = value_norm)) +
  geom_hline(yintercept = 0, col = "black", linetype = "solid") +
  geom_quasirandom(col = "darkgrey") +
  geom_boxplot(fill = NA, col = "black", outlier.shape = NA) +
  facet_grid(key ~ target + replicate, scales = "free_y") +
  xlab("") +
  ylab("Intensity (A.U.)") +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


# Also, remove outliers from plot
ylimits <- tib_gather %>%
  group_by(key) %>%
  dplyr::summarise(q95 = quantile(value_norm, 0.95) * 1.3)
ylimits_dapi <- ylimits$q95[ylimits$key == "dapi"]
ylimits_gfp <- ylimits$q95[ylimits$key == "gfp"]
ylimits_ki67 <- ylimits$q95[ylimits$key == "ki67"]

tib_gather %>%  
  mutate(shape = case_when(key == "dapi" & value_norm > ylimits_dapi ~ T,
                           key == "gfp" & value_norm > ylimits_gfp ~ T,
                           key == "ki67" & value_norm > ylimits_ki67 ~ T,
                           T ~ F),
         value_norm = case_when(key == "dapi" & shape ~ ylimits_dapi,
                           key == "gfp" & shape ~ ylimits_gfp,
                           key == "ki67" & shape ~ ylimits_ki67,
                           T ~ value_norm)) %>%
  ggplot(aes(x = condition, y = value_norm, col = replicate)) +
  geom_hline(yintercept = 0, col = "black", linetype = "solid") +
  geom_quasirandom(aes(shape = shape)) +
  geom_boxplot(fill = NA, col = "black", outlier.shape = NA) +
  facet_grid(key ~ target, scales = "free_y") +
  xlab("") +
  ylab("Intensity (A.U.)") +
  scale_color_brewer(palette = "Accent") +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


### Conclusion

Several observations:

  * Even though this analysis is only 2D-based, it convincingly shows that IAA
    strongly depletes Ki67. Never to the extend of the loss of GFP signal.
  * Both antibodies are very similar in their extend of depletion. This effect 
    size was also seen for the Abcam antibody.
  
Overall, this is a positive result and I should generate sequencing libraries
to determine the effect of Ki67 knockdown on the two new antibodies. So far, so
good.


### SessionInfo

```{r sessionInfo}

sessionInfo()

```














